Capitalized Title Here

### 2021-02-02 CMPSpaper_writing/CMPSpaper_writing.Rmd
combined intro and background

The main task of firearm examination is to determine whether two bullets were fired from the same gun barrel or not. It has been playing an important role in terms of convictions in the United States criminal justice system. However, in recent years people became more cautious about firearms identification, and criticism of its subjective methodology arose. A report published by the National Academy of Sciences in 2009 states that "[m]uch forensic evidence-including, for example, bite marks and firearm and toolmark identification—is introduced in criminal trials without any meaningful scientific validation, determination of error rates, or reliability testing to explain the limits of the discipline." This is why we need to develop frameworks and algorithms that can bring an objective and quantitative assessment into the field.

The congruent matching profile segment (CMPS) method is one of those algorithms developed by NIST [cite?]. It is ... [briefly talk about the CMPS method / algorithm, provide details in the Implementation section]

However, no open-source implementation.

In this paper, we present an open-source implementation of the CMPS algorithm.

Introduce our methodology - with 3D scans, digital microscope, data and algorithms

A short example to illustrate how one can use this CMPS package to obtain a comparison

* 2 lands - the same source; 1 land from another; signatures; CMPS score;

outline what is going to come

## Background

@@

The congruent matching profile segment (CMPS) method is used for "objective comparison of striated tool marks" [original paper], and in our cases it is used for bullet signatures. In order to obtain bullet signatures from the x3p files of bullet land impressions, we used `R` package `x3ptools`[cite?] and `bulletxtrctr`[cite?] with an automated framework proposed in [AUTOMATIC MATCHING OF BULLET LAND IMPRESSIONS]. `x3ptools` can be used to process x3p files. In the case of Hamby set 252, we need to convert the measurements of x3p files from meters to microns (`x3ptools::x3p_m_to_mum`) and rotate x3p files to make the long axis horizontal (`x3ptools::x3p_rotate`). Figure x.a shows a bullet land impression scan of a processed x3p file. The raised portion of the surface are groove impressions, and the vertical striation marks between the groove impressions are what we are interested in, called a land impression. The first step of obtaining the bullet signature is to extract a cross-section at a fixed height along the land impression. `bulletxtrctr::x3p_crosscut_optimize` can be used to determine an ideal height, and `bulletxtrctr::x3p_crosscut` can be used to acquire the crosscut data (`ccdata`). Figure x shows the typical pattern of the crosscut data, where "wedges" can be found on the left and right side and indicate the boundary between the land impression and the groove impression. `bulletxtrctr::cc_locate_grooves` is used to locate and chop these wedge areas so that only the curve of the land impression in the middle is kept. With the groove impressions excluded, we then use `bulletxtrctr::cc_get_signature` to smooth out the land impression curvature using LOESS (locally estimated scatter plot smoothing) [Cleveland 1979?] and can obtain the bullet signature. Figure x illustrates this framework. Sometimes the algorithm in our framework cannot determine the ideal location of the crosscut or the grooves due to the deformation of a bullet, so we also developed a `R` Shiny App called `bulletinspectR` to identify and correct those errors.

[Note that in the NIST CMPS paper they have another way of getting bullet signatures]

@@

```{r}
library(CMPS)
data(bullets)
```

>>>>

The main task of firearm examination is to determine whether two bullets were fired from the same gun barrel or not. It is achievable because lands in a gun barrel create striation marks on the bullet during the firing process, and it is believed that the same gun barrel will create highly similar striation marks on the bullets it fired. Typically the striation marks are created by lands in a barrel, so the striation marks are also sometimes refered as land impressions, and the areas the marks occupy are called land engraved areas (LEA). Firearm examination has been playing an important role in terms of convictions in the United States criminal justice system. However, in recent years people became more cautious about firearms identification, and criticism of its subjective methodology arose. A report published by the National Academy of Sciences in 2009 states that "[m]uch forensic evidence-including, for example, bite marks and firearm and toolmark identification—is introduced in criminal trials without any meaningful scientific validation, determination of error rates, or reliability testing to explain the limits of the discipline." To overcome those critics and concerns, people have been making an effort to build databases and develop frameworks and algorithms that can bring an objective and quantitative assessment into the field. The bullet signatures extracted from the bullet are important in this process since they can well describe the striation marks on the bullet and can well be understood by algorithms.

@@

The automated framework we used in this paper was proposed by Hare and Hofmann in 2017 [AUTOMATIC MATCHING OF BULLET LAND IMPRESSIONS?]. In order to obtain bullet signatures from the x3p files of bullet land impressions, we used `R` packages `x3ptools`[cite?] and `bulletxtrctr`[cite?]. `x3ptools` can be used to process x3p files. In the case of Hamby set 252, we need to convert the measurements of x3p files from meters to microns (`x3ptools::x3p_m_to_mum`) and rotate x3p files to make the long axis horizontal (`x3ptools::x3p_rotate`). Figure x.a shows a bullet land impression scan of a processed x3p file. The raised portion of the surface are groove impressions, and the vertical striation marks between the groove impressions are what we are interested in, called a land impression. The first step of obtaining the bullet signature is to extract a cross-section at a fixed height along the land impression. `bulletxtrctr::x3p_crosscut_optimize` can be used to determine an ideal height, and `bulletxtrctr::x3p_crosscut` can be used to acquire the crosscut data (`ccdata`). Figure x shows the typical pattern of the crosscut data, where "wedges" can be found on the left and right side and indicate the boundary between the land impression and the groove impression. `bulletxtrctr::cc_locate_grooves` is used to locate and chop these wedge areas so that only the curve of the land impression in the middle is kept. With the groove impressions excluded, we then use `bulletxtrctr::cc_get_signature` to smooth out the land impression curvature using LOESS (locally estimated scatter plot smoothing) [Cleveland 1979?] and can obtain the bullet signature. Figure x illustrates this framework. Sometimes the algorithm in our framework cannot determine the ideal location of the crosscut or the grooves due to the deformation of a bullet, so we also developed a `R` Shiny App called `bulletinspectR` to identify and correct those errors. Note that NIST researchers used a different way of getting the bullet signatures in their paper [cite?].

Once we obtain the bullet signatures, we can compare them using an algorithm that computes a similarity score. The congruent matching profile segment (CMPS) algorithm, developed by NIST for "objective comparison of striated tool marks" [original paper] in 2019, is one such algorithm. The main idea of the CMPS algorithm is that, we cut the reference profile (one of the two bullet signatures) into consecutive and non-overlapping basis segments and ask each basis segment to find the "best" registration position on the comparison profile (the other bullet signature) in terms of correlation values. Then the congruent registration position is the position where the most basis segments find it the "best", and the number of basis segments taking the congruent registration position is the CMPS score. The CMPS algorithm can assist firearm examiners with drawing a conclusion, but NIST's implementation of the CMPS algorithm was not released for the public. Thus, in this paper we present an open-source implementation of the CMPS algorithm so that everyone can use and improve it. With our implementation, one can compute the CMPS score of a comparison using the following code:

```{r, eval=FALSE}
cmps.result <- extract_feature_cmps(sig1, sig2)
```

In the example in section "Implementation" we would see that, two bullet signatures coming from the same source get a CMPS score of X, while two signatures coming from different sources get a CMPS score of Y. We also implemented graphing tools for users to better understand these results as well as the algorithm itself.

The section "Implementation" will go through the algorithm and show how to use the "CMPS" package. An example that illustrates the main points are also included. The section "Results" presents the results of evaluating the CMPS package using Hamby set 252 and Hamby set 44. And the last section covers some final discussion and conclusions.

@@

```{r, eval=FALSE}
library(CMPS)
data(bullets)
```
